\documentclass{book}
\input{preamble.tex}
\input{macros.tex}

\title{\Huge{MATH 105 Notes}\\\emph{Professor: Khalilah Beal-Uribe}\\\emph{Book: Real Mathematical
Analysis\thanks{An introductory but holistic, intuitive, and easy to read book.} by Pugh}}
\author{\huge{Neo Lee}}

\date{\huge{Spring 2024}}

\begin{document}

\maketitle
\let\cleardoublepage\clearpage
\pdfbookmark[section]{\contentsname}{toc}
\tableofcontents

\chapter{First chapter}
\section{Lecture 1}
\dfn{Norm}{
	Given a vector space $V$ over a subfield $\F$ of $\C$, a norm of $V$ is a real-valued function
	$p:V\to\R$ satisfying the following properties:
	\begin{enumerate}
		\item \tb{Triangle inequality:} $p(v+w)\leq p(v)+p(w)$,
		\item \tb{Absolute homogeneity:} $p(\alpha v) = |\alpha|p(v)$,
		\item \tb{Positive definiteness:} $p(v) \geq 0$ and $p(v) = 0$ iff $v = 0$.
	\end{enumerate}
	\nt{
		Usually, we denote the norm of $v$ by $\norm{v}$, and for clarity of the underlying vector 
		space, we may write $\norm{v}_V$.
	}
}
\mprop{Normed space is a metric space}{
	Let $V$ be a normed space. Then the function $d:V\times V\to\R$ defined by
	$d(v,w) = p(v-w) = \norm{v-w}$ is a metric on $V$.
}
\dfn{Isomorphism in vector spaces}{
	A function $f:V\to W$ between two vector spaces $V$ and $W$ over the same field $\F$ is called an
	isomorphism if it is bijective and linear. If such an isomorphism exists, we say that the two vector
	spaces are isomorphic.	
}
\dfn{Homeomorphism}{
	A function $f:X\to Y$ between two topological spaces $X$ and $Y$ is called a homeomorphism if it
	satisfies the following properties:
	\begin{enumerate}
		\item $f$ is bijective,
		\item $f$ is continuous,
		\item $f^{-1}$ is continuous.
	\end{enumerate}
	If such a homeomorphism exists, we say that the two topological spaces are homeomorphic.
}
\nt{
	In general, isomorphism does not imply homeomorphism. However, in certain cases, they are
	equivalent, which will be discussed in details later.
}
\dfn{Operator norm}{
	Let $T:V\to W$ be a linear oepration between normed spaces. Denote $\norm{\cdot}_V$ and
	$\norm{\cdot}_W$ be the norms in $V$ and $W$ respectively. The operator norm of $A$ is defined 
	by 
	\begin{align*}
		\norm{T} & = \sup \l\{\frac{\norm{Tv}_W}{\norm{v}_V}: v\neq 0, v\in V \r\} \\
		& = \inf \l\{c\geq 0: \norm{Tv}_W\leq c\norm{v}_V, \forall v\in V \r\}
	\end{align*}
	\nt{
		We say that $T$ is bounded if $\norm{T} < \infty$.
	}
}

\section{Lecture 2}
\thm{Multiplication of matrices are composition of linear maps}{
	$T_A\circ T_b=T_{AB}.$
}
\thm{Bounded operator is equivalent to continuous}{
	Let $T:V\to W$ be a linear transformation from one normed space to another. The following are
	equivalent:
	\begin{enumerate}
		\item $\norm{T} < \infty$,
		\item $T$ is uniformly continuous,
		\item $T$ is continuous,
		\item $T$ is continuous at $0$.
	\end{enumerate}
	\pf{Proof}{
		We show that $(1)\implies (2)\implies (3)\implies (4)\implies (1)$.
		\begin{itemize}
			\item $(1)\implies (2)$: Let $M=\norm{T}<\infty$, and let $\delta = \frac{\epsilon}{M}$.
			Then for any $x,y\in V$ such that $\norm{x-y}<\delta$, we have 
			\begin{align*}
				\norm{Tx-Ty} & = \norm{T(x-y)} \\
				& \leq M\norm{x-y} \\
				& < M\delta \\
				& = \epsilon.
			\end{align*}
			Hence, $T$ is uniformly continuous.

			\item $(2)\implies (3)$: Trivial. Uniformly continuous automatically implies continuous.
			\item $(3)\implies (4)$: Trivial. $T$ is continuous over the whole domain implies that
			it is continuous at any point in the domain, including $0$.

			\item $(4)\implies (1)$: Let $\epsilon=1$, then there exists $\delta>0$ such that 
			$\norm{x}<\delta$ implies $\norm{Tx}<1$. Then for any $v\neq 0$, define
			$v'=\frac{\delta}{2\norm{v}}$, then $\norm{v'}<\delta$ and hence
			$\norm{Tv'}<1$. Then we have 
			\begin{align*}
				\norm{Tv'} & < 1\\
				\l\|T\l(\frac{\delta}{2\norm{v}}v\r)\r\| & < 1 \\
				\frac{\delta}{2\norm{v}}\norm{Tv} & < 1 \\
				\norm{Tv} & < \frac{2}{\delta}\norm{v}.
			\end{align*}
			Then, from our \emph{definition 1.1.4} of operator norm, we have
			$\norm{T}<\frac{2}{\delta}$ and hence $\norm{T}<\infty$.
		\end{itemize}
	}
}
\thm{Linear map from finite-dimensional Euclidean space to normed space is continuous}{
	Let $T:\R^n\to W$, where $T$ is linear and $W$ is a normed space. Then 
	\begin{enumerate}
		\item $T$ is continuous,
		\item if $T$ is an isomorphism, then $T$ is a homeomorphism.
	\end{enumerate}
}
\cor{Linear maps from finite-dimensional normed space to normed space are continuous}{
	All linear maps from finite-dimensional normed space to another normed space are continuous, and all
	isomorphisms from finite-dimensional space to normed space are homeomorphisms.

	In particular, if a finite-dimensional vector spaces is equipped with two norms, then the
	identity map between them is a homeomorphism. For example, $T:\M\to\L$ is a homeomorphism.
	\pf{Proof}{
		Let $V$ be a $n$-dimensional normed space and $W$ be another normed space, and $T:V\to W$. Then,
		there exists an isoemorphism $S:V\to\R^n$. \emph{Theorem 1.2.2} gaurentees that $S$ and $S^{-1}$
		are homeomorphisms. Then, $T\circ S:\R^n\to W$ is also a continuous linear map guaranteed by
		\emph{Theorem 1.2.2}. Then, $$T = (T\circ S)\circ S^{-1}$$ is also a continuous linear because
		it is a composition of continuous linear maps. Hence, $T$ is continuous.

		Now, if $T:V\to W$ is an isomorphism where $V$ is a finite-dimensional normed space. Then, 
		$W$ is also a finite-dimensional normed space. Then, $T$ is continuous by the above argument.
		Then, $T^{-1}:W\to V$ is a linear map from a finite-dimensional normed space, hence also 
		continuous. Therefore, $T$ is a homeomorphism.

		Finally, let $V$ be a finite-dimenisonal vector space equipped with two norms
		$\norm{\cdot}_1$ and $\norm{\cdot}_2$. Then, the identity map $I:V\to V$ is an isomorphism 
		between the two finite-dimensional normed spaces. Then, $I$ is a homeomorphism by the above 
		argument.
	}
}

\section{Lecture 3}
Our goal is to generalize on-variable differentiation to multi-variable differentiation. In more
precise terms, we want to:
\nt{
	Obtain a natural derivative of $F:U\to\R^m$ at a point $p\in$ open set $U\subseteq\R^n$ by
	generalizing the derivative of $f:U\to \R$ at a point $p\in U\subseteq\R$.

	The key is to understand that $f$ is differentiable at $p$ if and only if $f$ is "approximately
	linear" at $p$.
}
	Consider an example in 2-dimensional Euclidean space to motivate our new definition of
	derivatives in multi-dimensional spaces. 
\ex{Derivative in $\R^2$}{
	Is $f:\R^2\to\R^2$ defined by $$f(x_1,x_2):=(x_1^2,x_2^2)$$ differentiable at the point
	$(1,2)\in\R^2$? 

	\sol{
		Let's first try to use the definition of derivative in $\R$ to see if it works. Let
		$p=(1,2)$, then we have 
		$$f'(p) = \lim_{h\to 0}\frac{f(p+h)-f(p)}{h} = \lim_{h\to
		0}\frac{f(\inner{1,2}+h)-f(1,2)}{h}$$ where $h$ is a vector. But this does not make sense 
		because we have not defined what it meant by division of a vector by a scalar. Hence, we
		need a new definition of derivative in $\R^2$, or more generally in multi-dimensional
		spaces. 
	}
}
\dfn{Multi-variable derivative (aka total derivative or Frechet derivative)}{
		Let $f:U\to\R^m$ be given where $U$ is an open subset of $\R^n$. The function $f$ is
		differentiable at $p\in U$ with derivative $(Df)_p=T$ if $T:\R^n\to\R^m$ is a linear
		transformation and 
		$$f(p+v)=f(p)+T(v)+R(v)\implies \lim_{\norm{v}\to 0}\frac{R(v)}{\norm{v}}=0.$$
		\nt{
			The form is coming from the definition of derivative in $\R$ by rearranging the terms in 
			$$f'(x)=\lim_{h\to0}\frac{f(x+h)-f(x)}{h}$$ to $$f(x+h)=f(x)+f'(x)h+R(h)\implies
			\lim_{h\to0}\frac{R(h)}{\norm{h}}=0.$$ We say that they Taylor remainder $R$ is
			\emph{sublinear} because it tends to 0 faster than $\norm{v}$.
		}
		\nt{
			Our definition of differentiability is coordinate free, which means we can study
			differentiation on spaces other than $\R^n$, e.g. differential manifolds, which is the
			natural next topic to study after $\R^n$.
		}
}
\ex{Back to \emph{example 1.3.1}}{
	Under our new definition, we can try to determine the differentiability of $f$ at $p=(1,2)$.
	Write \begin{align*}
		f(p+v) & = f(1+v_1,2+v_2) \\
		& = \inner{(1+v_1)^2,(2+v_2)^2} \\
		& = \inner{1+2v_1+v_1^2,4+4v_2+v_2^2} \\
		& = \inner{1,4} + \inner{2v_1,4v_2} + \inner{v_1^2,v_2^2} \\
		& = f(p) + \inner{2v_1,4v_2} + \inner{v_1^2,v_2^2}.
	\end{align*}
	Then, we can define $T:\R^2\to\R^2$ by $T(v_1,v_2)=\inner{2v_1,4v_2}$ and $R:\R^2\to\R^2$ by
	$R(v_1,v_2)=\inner{v_1^2,v_2^2}$. Now we just have to show that
	$\lim_{\norm{v}\to0}\frac{R(v)}{\norm{v}}=0$. We can check whether the norm of
	$\frac{R(v)}{\norm{v}}$ go to 0 when $\norm{v}\to0$. It doesn't matter which norm we choose (the
	Euclidean norm, sum norm, or max norm, etc.), because they are equivalent in finite-dimensional
	spaces. Let's choose the Euclidean norm for simplicity. Then, we have
	\begin{align*}
		\l\|\frac{R(v)}{\norm{v}}\r\| = \frac{1}{\norm{v}}\norm{\inner{v_1^2, v_2^2}} 
		& = \sqrt{\frac{v_1^4 + v_2^4}{v_1^2+v_2^2}} \\
		& = \sqrt{\frac{v_1^4}{v_1^2+v_2^2} + \frac{v_2^4}{v_1^2+v_2^2}} \\
		& \leq \sqrt{\frac{v_1^4}{v_1^2} + \frac{v_2^4}{v_2^2}} = \sqrt{v_1^2 + v_2^2} = \norm{v},
	\end{align*}
	which obviously goes to 0 when $\norm{v}\to0$. Hence, $f$ is differentiable at $p=(1,2)$ with 
	derivative $T(v_1,v_2)=\inner{2v_1,4v_2}$.
	\nt{
		Note: this $T$ is only true for $p=(1,2)$. For other points, we may have different derivatives. 
	}
}
\thm{Derivative is unique}{
	If $f$ is differentiable at $p$, then it uniquely determins $(Df)_p$ according to the limit
	formula, valid for all $u\in\R^n$,
	$$(Df)_p(u)=\lim_{t\to0}\frac{f(p+tu)-f(p)}{t}.$$
	\pf{Proof}{
		Let $T$ be a linear transformation that satisfies \emph{definition 1.3.1}. Now fix any
		$u\in\R^n$ and take $v=tu$. Then
		$$\frac{f(p+tu)-f(p)}{t}=\frac{T(tu)+R(tu)}{t}=T(u)+\frac{R(tu)}{t\norm{u}}\norm{u} = T(u) +
		\frac{R(tu)}{\norm{tu}}\norm{u}.$$ The last term converges to 0 as $t\to 0$ since
		$\norm{tu}\to 0$. Limits, when they exist, are unique, so $T(u)$ is uniquely determined.
	}
}
\thm{Differentiability implies continuity}{
	\pf{Proof}{
		Differentiability at $p$ implies that $$|f(p+v)-f(p)|=|(Df)_p(v)+R(v)|\leq |(Df)_p(v)|+|R(v)|
		= \norm{(Df)_p}|v| + |R(v)|,$$ which tends to 0 as $v\to 0$ since $\norm{(Df)_p}$ is finite
		in a finite-dimensional space and $R$ is sublinear. Hence, $f$ is continuous at $p$.
	}
}
\cor{Total derivative existsnce implies partial derivative existence}{}
\ex{Some important concepts}{
	\begin{itemize}
		\item All partial derivatives exist at a point does not imply total derivative exists at
		that point.
		\item All directional derivatives exist at a point does not imply total derivative exists at
		that point.
		\item Partial derivatives exist and are continuous at a point implies total derivative
		exists at that point.
	\end{itemize}
}


\chapter{Starting a new chapter}
\section{Demo of commands}
\dfn{Some defintion}{
	yap
}
\qs{Some question}{yap}
\sol{
	\pf{Some proof}{yap}
}
\nt{Some note}
\thm{Some theorem}{
	yap
}
\wc{Some wrong concept}{
	yap
}
\mlemma{Some lemma}{
	yap
}
\mprop{Some proposition}{
	yap
}
\ex{Some example}{
	yap
}
\clm{Some claim}{
	yap
}
\cor{Some corollary}{
	yap
}
\thmcon{Some unlabeled theorem}

This is a new paragraph


\begin{algorithm}
	\caption{Some algorithm}
	\KwIn{input}
	\KwOut{output}
	\SetAlgoLined
	\SetNoFillComment
	\tcc{This is a comment}
	This is first line \tcp*{This is also a comment}
	\uIf{$x > 5$} {
		do nothing
	} \uElseIf {$x < 5$} {
		do nothing
	} \Else {
		do nothing
	}
	\While{$x == 5$}{
		still do nothing
	}
	\ForEach{$x = 1:5$}{
		do nothing
	}
	\Return{return nothing}
\end{algorithm}


\end{document}
