\documentclass{book}
\input{preamble.tex}
\input{macros.tex}

\title{\Huge{MATH 185 Notes}\\\emph{Professor: Constantin Teleman}}
\author{\huge{Neo Lee}}
\date{\huge{Spring 2024}}

\begin{document}

\maketitle
\let\cleardoublepage\clearpage
\pdfbookmark[section]{\contentsname}{toc}
\tableofcontents

\chapter{First chapter}
\section{Lecture 1 (skipped)}
\section{Lecture 2}
\dfn{Integral powers of complex numbers from geomtery}{
	If $z=r\cdot e^{i\theta}$, so $\abs{z}=r$ and define $\arg{z}=\theta$, 
	then $z^n=r^n\cdot e^{in\theta}$.
	\nt{
		$\abs{z} = r$ can be shown by writing $e^{i\theta}=\cos\theta+i\sin\theta$ and taking absolute
		value. 
	}
	\nt{
		Because of our definition, integral powers of complex numbers can be easily visualized in polar
		coordinates. In particular, for $z=r\cdot e^{i\theta}$, $$\abs{z}=r^n\quad \t{and} \quad
		\arg{z^n} = n\theta.$$
	}
}
\ex{Map from $z\to z^2$}{
	Let $z=x+yi$, then $\re z$ is mapped from $x$ to $x^2$, while $\im z$ is mapped from $yi$ to
	$-y^2$.

	$$z=x+yi\mapsto x^2-y^2+2xyi.$$ Hence, $\re z^2 = x^2-y^2$ and $\im z = 2xy$. If we plot $z^2$
	on the real and imaginary axis, then it's a parabola.
}

\thm{De Moivre formula}{
	$$(\cos\theta+i\sin\theta)^n = \cos(n\theta)+i\sin(n\theta).$$
	\pf{Proof}{
		This falls immediately by taking $n^\t{}th$ power of $z=e^i\theta$ and following
		\emph{definition 1.1}.
	}
	\nt{
		By applying binomial expansion to the left hand side of De Moivre formula, we get the following
		consequences immediately:
		\begin{align*}
			\cos^n\theta - {n\choose2} \cos^{n-2}\theta\cdot\sin^2\theta +\cdots & = \cos(n\theta) \\
			i\l(n\cdot\cos^{n-1}\theta\cdot\sin\theta - {n\choose3}\cos^{n-3}\theta\cdot\sin^3\theta + \cdots \r) & = i\sin(n\theta).
		\end{align*}
		From the first equation, we can see $\cos^2\theta - \sin^2\theta = \cos(2\theta)$ is a
		special case of De Moivre formula. From the second equation, we can see
		$2\cos\theta\cdot\sin\theta = \sin(2\theta)$ is a special case as well.
	}
}

\dfn{$n^\text{th}$ roots of complex number}{
	The $n^\t{th}$ roots of complex number are the solutions $z$ to the equation $z^n=x$.
	\nt{
		We call the solutions $z$ to the equation $z^n=1$ the $n^\t{th}$ roots of unity.
	}
}

\mprop{There are $n^\t{th}$ distinct roots of unity}{
	There are $n^\t{th}$ distinct roots $z$ to the equation $z^n=1$, and they are 
	$$1,\cos\frac{2\pi}{n}+i\sin\frac{2\pi}{n}, \cos\frac{4\pi}{n}+i\sin\frac{4\pi}{n},\cdots,
	\cos\frac{(2n-2)\pi}{n} + i\sin\frac{(2n-2)\pi}{n}.$$
	Let $\xi_n = \cos\frac{2\pi}{n} + i\sin\frac{2\pi}{n}$, then the list is 
	$$1,\xi_n, \xi_n^2, \dots, \xi_n^{n-1}.$$
	\nt{
		The roots are essentially cutting the unit circle into $n$ parts, and the cuts lying on the unit
		circle are the roots.
	}
}

\cor{Linear factorization of $x^n-1$}{
	From the previous proposition, we have the linear factorization 
	$$x^n-1=(x-1)(x-\xi_n)(x-\xi_n^2)\cdots(x-\xi_n^{n-1}).$$
	\pf{Proof}{
		From Bezout's theorem, if $x=a$ is a solution to $P(x)=0$, then $(x-a)$ divides $P(x)$. Now
		from \emph{proposition 1.1}, we found all the roots to $x^n-1=0$, so we can linearly factor
		$x^n-1$. 
	}
}

\dfn{Primitive roots of unity}{
	An $n^\t{th}$ root of unity $\xi$ is called primitive if $\xi^n=1$ but $\xi^k\neq 1$ for all
	$0<k<n$. In other words $\xi$ is the $n^\t{th}$ primitive root of unity if $\xi$ raised to the
	power $n$ is the first time equating to 1.
}

\ex{$6^\t{th}$ roots of unity}{
	Consider the equation $z^6=1$ and let $\xi_n = \cos\frac{\pi}{3}+i\sin\frac{\pi}{3}$. From
	\emph{proposition 1.1}, we know all the roots are raising $\xi$ to the power until $5$. Then by
	drawing it geometrically as points on the unit circle, and raising power is equavalent to
	rotating along the circle, the $6^\t{th}$ primitive roots are $\xi$ and $\xi^5$.
}

\thm{$n^\t{th}$ roots of unity can be partition into divisors of $n$}{
	Let $\{d_1, \dots, d_k\}$ be the divisors of $n$, then the set of $n^\t{th}$ roots of unity
	can be partitioned into groups of primitive $d_k^\t{th}$ roots of unity.
}

\thm{Factorization of $x^n-1$ in integer coefficient terms}{
	Define the $d^\t{th}$ cyclotomic polynomial as $\Phi_d(x)$, which is an irreducible polynomial
	over rationals and with integer coefficients. Then, grouping the factors of
	$$x^n-1=(x-1)(x-\xi_n)\cdots(x-\xi_n^{n-1})$$ according to divisors of $n$ with primitive roots
	gives a factorization $$x^n-1=\prod_{d\vert n}\Phi_d(x).$$
}

\ex{Factorization of $x^3-1$ and $x^6-1$}{
	\begin{align*}
		x^3-1 & = \underbrace{(x-1)}_{\Phi_1(x)}\underbrace{(x^2+x+1)}_{\Phi_3(x)} \\
		x^6-1 & = \underbrace{(x-1)}_{\Phi_1(x)}\underbrace{(x+1)}_{\Phi_2(x)}
		\underbrace{(x^2+x+1)}_{\Phi_3(x)}\underbrace{(x^2-x+1)}_{\Phi_6(x)}.
	\end{align*}
}

\dfn{"Cyclotomic integers"}{
	Cyclotomic integers are $\Z[\xi_2,\xi_3,\dots] = \Z[\xi_n]$ for a fixed $n$. Equivalently, it's
	formed by adjoining all roots of units to $\Z$.
}

\section{Lecture 3 (skipped)}
\section{Lecture 4}
\subsection{Multi-variable real differentiation}
\nt{
	Naturally constructed functions will often have singularities, for example, the inversion 
	$$z\mapsto \frac{1}{z}.$$ As $z\to0$ in \C, $\abs{z}\to0$ in \R, so
	$\abs{z^{-1}}=\abs{z}^{-1}\to\infty$. This can be fixed by the extended complex plane $\C\cup
	\{\infty\}.$ 
}
\dfn{Real differentiation in $\R$}{
	Let $I\in\R$ be an open interval. A function $f:I\to\R$ is differentiable at a point $x_0\in I$
	if $$\lim_{x\to x_0}\frac{f(x)-f(x_0)}{x-x_0}$$ exists.
}
\ex{Motivating example of new defintion of $\R^2$ differentiation}{
	We say $f(x_0)+f'(x_0)(x-x_0)$ is the best linear appriximation to $f$ at point $x_0$, and 
	$$\frac{|f(x)-f(x_0)-f'(x_0)(x-x_0)|}{|x-x_0|}\to0\quad\t{as}\quad x\to x_0.$$
	\nt{
		With some knowledge of multivariable analysis, we can get this definition by rearranging the
		definition of multi-variable differentiation involving the Taylor remainder $R(x-x_0)$.
	}
}
\dfn{Real differentiation from $\R^2\to\R$}{
	Let $U\in\R^2$ be an open subset. A function $f:U\to\R$ is differentiable at $\inner{x_0,y_0}\in
	U$ if there exists a linear map $Df:\R^2\to\R$ such that 
	$$\frac{|f(\inner{x,y})-f(\inner{x_0,y_0})-Df(\inner{x-x_0,y-y_0})|}{\norm{\inner{x-x_0,y-y_0}}}\to
	0\quad\t{as}\quad\inner{x,y}\to\inner{x_0,y_0}.$$ 
	\nt{
		$Df$ is the derivative at $\inner{x_0,y_0}$, which gives the best linear approximation to $f$ 
		at $\inner{x_0,y_0}$.
	}
}
\dfn{Partial differentiation from $\R^2\to \R$}{
	Let $f:U\to\R$ where $U$ is an open subset of $\R^2$, then $$\pdv[]{f}{x}=\lim_{x\to
	x_0}\frac{f(\inner{x,y_0})-f(\inner{x_0,y_0})}{x-x_0}\quad\t{and}\quad\pdv[]{f}{y}=\lim_{x\to
	x_0}\frac{f(\inner{x_0,y})-f(\inner{x_0,y_0})}{y-y_0}.$$
	\nt{
	\begin{itemize}
		\item 
		This is reducible to the definition of real differentiation in $\R$ by fixing $x=x_0$ or
		$y=y_0$.

		\item
		$Df$ as the derivative at $\inner{x_0,y_0}$ in this case is given by the (matrix)
		multiplication by the vector $\inner{\pdv[]{f}{x}, \pdv[]{f}{y}}$.
	\end{itemize}	
	}
}
\dfn{Differentiation from $\R^2$ to $\R^2$}{
	Obvious adoptation of the previous definition, where $f$ is a vector valued function, say
	$\R^2\to\R^2$. If $f=\begin{bmatrix}
		f_1 \\ f_2
	\end{bmatrix}$, $Df$ is the Jacobian matric $\begin{bmatrix}
		\pdv[]{f_1}{x} & \pdv[]{f_1}{y} \\
		\pdv[]{f_2}{x} & \pdv[]{f_2}{y}
	\end{bmatrix}$.
}
\thm{Sufficient condition for total differentiability}{
	If all partial derivatives of $f$ exist and are continuous in a neighborhood of
	$\inner{x_0,y_0}$, then $f$ is totally differentiable at $\inner{x_0,y_0}$.
}
\nt{
	For limits and convergence questions in \C, we can consider it as limits and convergence in
	$\R^2$. However, note that we cannot treat differentiability in $\C$ directly as
	differentiability in $\R^2$, which we will later prove that there are extra conditions.
}
\nt{
	Addition and multiplication of complexnumbesr are continuous functions $\C\times\C\to\C$, and
	inversion map $z\to z^{-1}=\frac{z}{\abs{z}^2}$ is also continuous away from $z=0$.
}

\subsection{Complex differentiation}
\dfn{Complex differentiable}{
	$f$ is complex differentiable at $z_0$ if $$\lim_{z\to z_0}\frac{f(z)-f(z_0)}{z-z_0}$$ exists.
	\nt{ 
		The limit is the (complex) derivative of $f$ at $z_0, f'(z_0)$.
	}
}
\thm{Complex differentiability implies continuity}{
	If $f$ is complex differentiable at $x_0$, then $f$ is continuous at $z_0$.
	\pf{Proof}{
		\begin{align*}
			\lim_{z\to z_0}f(x)-f(x_0) & = \lim_{z\to z_0}\l(f(x)-f(x_0)\r)\frac{z-z_0}{z-z_0} \\
			& = \lim_{z\to z_0}\frac{f(z)-f(z_0)}{z-z_0} \cdot \lim_{z\to z_0}z-z_0 \\
			& = f'(z_0)\cdot 0 = 0.
		\end{align*}
	}
}
\ex{Derivative of monomial}{
	Let $f(z)=z^n$, then notice $$\frac{f(z)-f(z_0)}{z-z_0}=\frac{z^n-z_0^n}{z-z_0}=z^{n-1}+z^{n-2}z_0
	+\cdots+z_0^{n-1}.$$ Since addition and multiplication are continuous on $\C$ and $\lim_{z\to z_0}z =
	z_0$, the limit of the RHS is $nz_0^{n-1}$. Hence, $f'(z_0)=nz_0^{n-1}$.
}
\thm{Properties of complex differentiation}{
	If $f,g$ are complex differentiable at $z_0$, then
	\begin{enumerate}
		\item $f+g$ is complex differentiable at $z_0$ and $(f+g)'(z_0)=f'(z_0)+g'(z_0)$.
		\item $fg$ is complex differentiable at $z_0$ and $(fg)'(z_0)=f'(z_0)g(z_0)+f(z_0)g'(z_0)$.
		\item If $g(z_0)\neq 0$, then $\frac{f}{g}$ is complex differentiable at $z_0$ and
			$$\l(\frac{f}{g}\r)'(z_0)=\frac{f'(z_0)g(z_0)-f(z_0)g'(z_0)}{g(z_0)^2}.$$
	\end{enumerate}
	\pf{Proof}{
		The same proof as in $\R$ using absolute value of complex numbers and triangle inequality.
	}
}
\cor{Complex differentiable of rational functions}{
	Let 
	\[
		z = \frac{P(z)}{Q(z)} \quad\t{for some polynomials } P,Q \t{ and } Q \neq 0.
	\]
	We can omplex differentiate $z$ using the same algebraic manipulation as in $\R$. 
	\pf{Proof}{
		From \emph{example 1.4.2}, we know how to differentiate monomials just like in the real case,
		and from \emph{theorem 1.4.3}, we can differentiate polynomials just like in the real case by
		linear combination of monomials. Hence, we can differentiate $P(z)$ and $Q(z)$, and then apply
		\emph{theorem 1.4.3 (3)}.
	}
}
\ex{$f=\frac{1}{z}$}{
	$f'(z)=\frac{-1}{z^2}$ when $z\neq0$.
}
\dfn{Holomorphic}{
	$f:U\to\C$ is holomorphic if it is complex differentiable at every point in the open set
	$U\subseteq\C$.
}
\subsection{Complex vs real differentiability}
\mprop{Complex differentiabilit implies $\R^2$ differentiability}{
	Consider complex number $z=x+yi$ and $f(z)=u(x,y)+iv(x,y)$.
	If $f$ is complex differentiable at $z_0=x_0+y_0i$, then the map 
	\[
		\begin{bmatrix}
			u(x,y) \\ v(x,y)
		\end{bmatrix}: \R^2\to\R^2
	\]
	is real differentiable at $\inner{x_0,y_0}$.
	\pf{Proof}{
		If $f$ is differentiable at $z_0$, then 
		\[
			\lim_{z\to z_0}\frac{f(z)-f(z_0)-f'(z_0)(z-z_0)}{z-z_0} = 0,
		\]
		which is equivalent to 
		\[
			\lim_{z\to z_0}\frac{|f(z)-f(z_0)-f'(z_0)(z-z_0)|}{|z-z_0|} = 0.
		\]
		Now notice that in face $f'(z_0)$ is a linear homogeneous function from $\C\to\C$. Again if
		we consider $z_0=x_0+y_0i$, then $f'(z_0)$ is a linear function of $\inner{x_0,y_0}$ from
		$\R^2\to\R^2$. Hence, we have found the derivative $D\begin{bmatrix}
			u(x_0,y_0) \\ v(x_0,y_0)
		\end{bmatrix}$, which is $f'(z_0)$ when writing in vector $\R^2$ form.
		\nt{
			Since $f'(z_0)$ is a linear function from $\C\to\C$, it is simply scaling $z_0$ by a complex 
			number. if $f'(z_0) = (p + qi)z_0$, then $\re f'(z_0) = p\cdot\re z_0 - q\cdot\im z_0$ and
			$\im f'(z_0)=p\cdot\im z_0 + q\cdot\re z_0$. Writing it in matrix form and representing
			complex numbers as vectors in $\R^2$, we get the matrix 
			$
				\begin{bmatrix}
					p & -q \\
					q & p
				\end{bmatrix}.
			$
		}
	}	
}
\thm{Equivalence of complex and $\R^2$ differentiability}{
	\[
		(z=x+yi)\mapsto (f(z)=w=u(x,y)+v(x,y)i)
	\] is complex differentiable at $z_0 = x_0+y_0i$ if and only if the map 
	\[
		\begin{bmatrix}
			x \\ y
		\end{bmatrix}\mapsto 
		\begin{bmatrix}
			u(x,y) \\ v(x,y)
		\end{bmatrix}
	\] is real differentiable at $(x_0,y_0)$ AND satisfies the Cauchy-Reimann equations 
	\[
		u_x = v_y,\quad v_x = -u_y
	\] at $(x_0,y_0)$.
}
\ex{Exponential}{
	The complex exponential $f:z\mapsto e^z$ is holomorphic in \C.
	\pf{Proof}{
		\[
			f:z\mapsto e^z \iff f:(x+yi)\mapsto e^x(\cos y + i\sin y).	
		\]
		If we write it in $\R^2$, then 
		\[
			g:\begin{bmatrix}
				x \\ y
			\end{bmatrix}\mapsto 
			\begin{bmatrix}
				e^x\cos y \\ e^x\sin y
			\end{bmatrix}.
		\] We know the function is differentiable in $\R^2$ because the partial derivatives exist and 
		are continuous. In particular, the derivative is 
		\[
			Dg: \begin{bmatrix}
				x \\ y
			\end{bmatrix}\mapsto 
			\begin{bmatrix}
				e^x\cos y & -e^x\sin y \\
				e^x\sin y & e^x\cos y
			\end{bmatrix}.	
		\]
		Indeed, the Cauchy-Reimann equations are satisfied, and hence $f$ is real differentiable in
		$\R^2$, hence complex differentiable in \C, which is the definition of holomorphic.
	}
}
\ex{Other holomorphic / non-holomorphic examples}{
	\begin{itemize}
		\item $(x,y)\mapsto\l(x^2+y^2,2xy\r)$ \quad  non-holomorphic
		\item $(x,y)\mapsto\l(x^2-y^2,2xy\r)$
		\item $(x,y)\mapsto\l(\frac{1}{2}\log\l(x^2+y^2\r),\arctan\frac{y}{x}\r)$
	\end{itemize}
	\pf{Proof}{
		Conside as $U\to\R^2:U\subseteq\R^2$ map and check that the Jacobian matrix satisfies the 
		Cauchy-Reimann equations.
	}
}
\subsection{Some general properties of holomorphic functions}
\mprop{Set of holomorphic maps is an algebra}{
	The set of holomorphic maps $U\to\C$ is an algebra:
	\begin{itemize}
		\item If $f,g$ are holomorphic, then $f+g$ is holomorphic.
		\item If $k\in\C$, then $kf$ is holomorphic.
		\item If $f,g$ are holomorphic, then $fg$ is holomorphic.
		\item If $f,g$ are holomorphic and $g(z)\neq 0$ for all $z\in U$, then $\frac{f}{g}$ is
			holomorphic.
	\end{itemize}
}
\thm{Holomorphic function implies holomorphic derivative}{
	If $f:U\to\C$ is holomorphic and twice real differentiable, then $f'$ is also holomorphic.
	\nt{
		In fact, the assumption of twice real differentiable can be omitted because holomorphic maps are
		infinitely differentiable.
	}
	\pf{Proof}{
		Let $f(x+yi)=u(x,y)+v(x,y)i$ where $u,v:U\to\R$. Then the Jacobian matrix of $f$ is
		\[
			Df = \begin{bmatrix}
					u_x & u_y \\
					v_x & v_y
			\end{bmatrix} = 
			\begin{bmatrix}
				u_x & -v_x \\
				v_x & u_x
			\end{bmatrix}
		\]
		Or we can write $Df=(u_x-v_x) + (v_x+u_x)i$, then the Hessian matrix (derivative of $f'$) of
		$f$ is
		\[
			D^2f = \begin{bmatrix}
				u_{xx}-v_{xx} & u_{xy}-v_{xy} \\
				v_{xx}+u_{xx} & v_{xy}+u_{xy}
			\end{bmatrix} = 
			\begin{bmatrix}
				u_{xx}-v_{xx} & \overbrace{u_{yx} - v_{yx}}^{\t{Clairaut's}} \\
				u_{xx}+v_{xx} & \underbrace{v_{yx} + u_{yx}}_{\t{Clairaut's}}
			\end{bmatrix} = 
			\begin{bmatrix}
				u_{xx}-v_{xx} & \overbrace{-u_{xx} - v_{xx}}^{u_x = v_y,v_x = -u_y} \\
				u_{xx}+v_{xx} & \underbrace{u_{xx} - v_{xx}}_{u_x = v_y,v_x = -u_y}
			\end{bmatrix}.
		\]
		Hence, we have shown that for $D^2f=f''$ is also real differentiable satisfying the 
		Cauchy-Reimann equations. Hence, $f'$ is holomorphic.
	}
}






\chapter{Starting a new chapter}
\section{Demo of commands}
\dfn{Some defintion}{
	yap
}
\qs{Some question}{yap}
\sol{
	\pf{Some proof}{yap}
}
\nt{Some note}
\thm{Some theorem}{
	yap
}
\wc{Some wrong concept}{
	yap
}
\mlemma{Some lemma}{
	yap
}
\mprop{Some proposition}{
	yap
}
\ex{Some example}{
	yap
}
\clm{Some claim}{
	yap
}
\cor{Some corollary}{
	yap
}
\thmcon{Some unlabeled theorem}

This is a new paragraph


\begin{algorithm}
	\caption{Some algorithm}
	\KwIn{input}
	\KwOut{output}
	\SetAlgoLined
	\SetNoFillComment
	\tcc{This is a comment}
	This is first line \tcp*{This is also a comment}
	\uIf{$x > 5$} {
		do nothing
	} \uElseIf {$x < 5$} {
		do nothing
	} \Else {
		do nothing
	}
	\While{$x == 5$}{
		still do nothing
	}
	\ForEach{$x = 1:5$}{
		do nothing
	}
	\Return{return nothing}
\end{algorithm}


\end{document}
